Fair, Stable Exchange Networks through Quenched Merchant Location and Idiosyncratic Trading Costs

### 2019-06-24 writing/Paper/main.tex
Writing imporovements

\documentclass[11pt]{article}
\usepackage[a4paper,margin=2.8cm]{geometry}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage[dvipsnames,usenames]{color}
\usepackage[colorlinks=true,linkcolor=NavyBlue,citecolor=NavyBlue,urlcolor=NavyBlue]{hyperref} % for hyperlinks 
\usepackage{floatpag}

@@

With enough market power, a profit seeking firm will try to raise barriers to
entry to prevent competition. This stops firms that have new innovations from
surviving long enough to succeed in a market. The distance based tax introduced
in this paper allows start ups to overcome entrenched monopolies' unfair
advantages.

In this paper, we propose a new distance based tax, which we call an
idiosyncratic tax, in order to solve these
problems, then show that it works by modelling an economy that implements the
tax. In section 2 we give an overview of this economy. In section 3 we
formalize the model, and discuss how the firms compete on capacity and price.
In section 4, we discuss results of running the model. In section 5 we expand
the model by allowing for multiple timesteps, and having firms decrease their
costs based on previous profits as a form of investment.
%[change as we add sections]

\section{Economy}
We simulate an economy of buyers and firms in some environment with a distance
metric (e.g. a country, a network economy), then see how firms respond when we
apply the idiosyncratic tax.

@@

The left insets of Figure \pageref{fig:fig3x2SingleTimestep} are examples of
such a geography.

@@

\subsection{Idiosyncratic Tax}
The idiosyncratic tax that is introduced by this paper is dependent on a
distance measuere $D$, or ordinal distance, that gives a discrete distance
between every pair of buyers and firms. $D$ is defined such that the closest
firm to a buyer is at ordinal distance $D=1$, the next closest at $D=2$, and
the $n$th closest at $D=n$. Consider a firm and two buyers, one buyer near the
firm, the other distant. The idiosyncratic tax means that the nearer buyer can
buy more cheaply from that firm than can the more distant one. This
incentivizes buyers to purchase from more local firms.

@@

\section{Model Definition}
When firms choose capacity and price, they make their choices in order to
maximize profit. First, the firms compete on capacity, by simultaneously
declaring how much of the product they plan to produce; next, they compete on
price to determine the firms' non-tax prices. This process, consisting of these
two subgames, is called a Kreps-Scheinkman game. The major difference we
introduce to the classical Kreps-Scheinkman game is the idiosyncratic tax,
which significantly alters the pricing subgame.

@@

The capacity subgame acts much like a Cournot game.\footnote{A Cournot game is
where firms compete only on quantity, and sell all of the quantity at the
market clearing price.} The difference is that instead of being assigned a
market clearing price and selling everything they produce, in our model the
firms compete on price in a second subgame where they choose a price. Only
after that do the buyers determine how much of each firms capacity is sold.

\subsection{Pricing Game}
Each firm $i$, having chosen a capacity $\hat q_i$, now chooses a price, $p_i$.
Consider two vending machines right beside one another selling the same soda,
with one machine selling soda at a cheaper price than the other. In this
subgame, buyers will use the cheaper machine first, but if the buyers demand
more soda than that machine's capacity, buyers will switch to the more
expensive machine when the cheaper one runs out of soda. Competition on price
in such a manner is called an Edgeworth game. Edgeworth games are very
difficult to find an equilibrium in, as depending on demand and capacity, a
variety of strategies can be employed.

@@

Let the set of buyers be called $B$. Each buyer $b\in B$ buys a quantity $q_b =
\sum_{i\in S} q_{b,i}$, where $q_{b,i}$ is the amount buyer $b$ buys from the
firm $i$. The buyer's all have identical demand curves $p_\text{max} + q_b =
a$, where $a$ is the endowment they are given exogenously, and $p_\text{max}$
is the maximum price they pay (including tax). Each buyer tries to maximize
$q_b$ while keeping $P + Q \le a$. In other words, they buy from the cheapest
available firm until they hit the demand curve.

@@

\subsection{Profit Function}
Finally, firm $i$, having chosen quantity $\hat q_i$ and price $p_i$, and
having sold the product to buyers, will earn a profit:

@@

Note that while the firm can choose both their capacity $\hat q_i$ and their
price $p_i$, their cost per unit $c_i$ is set exogenously. Also, each $q_{b,i}$
is chosen by the buyer $b$ so that they buy as much as possible while staying
under their demand curve $P+Q\leq a$; this is where taxes affect the firm's
profit.

@@

\subsection{Numerical Approximation}
Because the buyers have individualized preferences and the pricing game does
not necessarily have a `best' strategy, it is difficult to determine a
generalized, theoretical solution to the model laid out above. Instead, we
numerically determine the Nash equilibrium.\footnote{A Nash equilibrium
for a game is a set of choices, one for each player, such that upon seeing the
other firm's choices, could increase their payoff by changing strategies. In
our model, the players are firms, the payoff is profit, and the choices are
either capacity or price, depending on the subgame.}

To find the capacities and prices that the firms will choose, we begin by
discretizing the capacity and pricing choices so we only have a finite number
of possible combinations. Then, for every outcome of the capacity game, we find
the Nash equilibrium of the pricing game using the Python library Gambit. Then,
using the equilibrium prices, we calculate the profits for every outcome of the
capacity game. We finally use this information to find the Nash equilibrium of
the capacity game, again using Gambit.

\section{Examples}
Figure \ref{fig:fig3x2SingleTimestep} contains the output of five instances
of finding the numerical solution for different $\gamma$ and costs. In the
first two rows $\gamma=0$, with the one difference being that in Row 1, both
Firm 1 and Firm 2 have the same cost, and in Row 2, Firm 1 has a 1\% lower cost
than Firm 2. In Row 1, the firms sell the same amount of product and make the
same amount of profit. But in Row 2, Firm 1 sells 33\% more quantity and makes
53\% more profit. As a side note, since when $\gamma = 0$ there is no tax on
distance, we can see on the left insets of these rows that buyers do not
necessarily buy from the closest firm.

In the Rows 3, 4, and 5, we let $\gamma = .5$ instead. In Row 3 we can see a
mirror of Row 1, where both firms, faced with the same costs, produce the same
quantity, and make the same profit. But in row 4, where we give Firm 1 the same
advantage we game them in the row 2, they sell the same amount, and only make
12\% more profit than Firm 2. To increase their market share, Firm 1 needs to
reduce their costs to 80\% of Firm 2, as shown in row 5.

The idiosyncratic tax thus acts as a dampener, preventing companies with
slight advantages from acquiring overwhelming market dominance, while still
ensuring that minor innovations are profitable. But, if one firm is
significantly better, as in Row 5, then that firm will still dominate the
market.

\begin{figure}[[p!htb] % p stands for 'entire page', '!htb' is to insert as close to code as possible
  \floatpagestyle{empty} % remove page number
  \centering
  \vspace{-2cm} % shift upwards to get more space for caption
  \includegraphics[width=\linewidth]{5x2plot.png}
  \caption{\small{Above are the model's results for a variety of
  conditions. The Rows 1 and 2 have $\gamma=0$, while Rows 3, 4, and 5 have
  $\gamma = .5$. In Rows 1 and 3, the firms have an equal cost per unit, sell
  equal quantity, and make equal profit. In Row 2, where $\gamma=0$, a 1\%
  difference in cost means Firm 1 has a 10\% greater market share, and a 53\%
  larger profit.
  In Row 4, where $\gamma=.5$, Firm 1's 1\% difference in cost now results in
  an equal market share, but they still profit 12\% more. To get a larger
  market share, as in Row 5, Firm 1 must have a 20\% lower cost per unit. This
  means that $\gamma$ acts as a dampener on monopolization by increasing the
  amount of innovation needed to gain market power while still encouraging
  innovation. %[Check numbers after new gamma]
  }}
  \label{fig:fig3x2SingleTimestep}
\end{figure}

\section{Dynamics}
We run this model once for every timestep, then update the costs for the firms
between the timesteps. Every timestep we determine which firm has made the most
profit. For that firm, we reduce its cost per unit by 5\% in order to simulate
the company investing its excess profits in itself to become more profitable in
the long term.

\end{document}

>>>>

\documentclass[11pt]{article}
\usepackage[a4paper,margin=2.8cm]{geometry}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage[dvipsnames,usenames]{color}
\usepackage[colorlinks=true,linkcolor=NavyBlue,citecolor=NavyBlue,urlcolor=NavyBlue]{hyperref} % for hyperlinks 
\usepackage{floatpag}
\usepackage{hyperref}
\usepackage[backend=biber,style=authoryear]{biblatex}
\bibliography{refs}

@@

\green{If we don't add the start up stuff, delete this.}
With enough market power, a profit seeking firm will try to raise barriers to
entry to prevent competition. This stops firms that have new innovations from
surviving long enough to succeed in a market. The distance based tax introduced
in this paper allows start ups to overcome entrenched monopolies' unfair
advantages.

In this paper, we propose an idiosyncratic tax based on distance in order to
solve these problems, then show that it works by modelling an economy that
implements the tax. In section 2 we give an overview of this economy. In
section 3 we formalize the model, and discuss how the firms compete on capacity
and price. In section 4, we discuss results of running the model. In section 5
we expand the model by allowing for multiple timesteps, and having firms
decrease their costs based on previous profits as a form of investment.
%[change as we add sections]

\section{Economy}
We simulate an economy of buyers and firms in some environment with a distance
metric (e.g. a country, a network economy), then observe how firms respond when
we apply the idiosyncratic tax.

@@

The left insets of Figure \ref{fig:fig3x2SingleTimestep} are examples of
such a geography.

@@

\subsection{Idiosyncratic Tax}
The idiosyncratic tax that is introduced by this paper is dependent on a
distance measure $D$, or ordinal distance, that gives a discrete distance
between every pair of buyers and firms. We define the ordinal distance $D$ be
defined such that the closest firm to a buyer is at ordinal distance $D=1$, the
next closest at $D=2$, and the $n$th closest at $D=n$. Consider a subset of a
larger economy consisting of a firm and two buyers, one buyer near the firm,
the other distant. \green{right here I am trying to express show that because
of the tax, two buyer's can perceive different prices at the same firm.} The
idiosyncratic tax means that the nearer buyer can buy more cheaply from that
firm than can the more distant one. This incentivizes buyers to purchase from
more local firms.

@@

\section{Model Definition}
When firms choose capacity and price, they make their choices in order to
maximize profit. First, the firms compete on capacity, by simultaneously
declaring how much of the product they plan to produce; next, they compete on
price to determine the firms' non-tax prices. This process, consisting of these
two subgames, is called a Kreps-Scheinkman game \cite{kreps1983quantity}.
The major difference we introduce to the classical Kreps-Scheinkman game is the
idiosyncratic tax, which significantly alters the pricing subgame.

@@

The capacity subgame acts much like a Cournot game.\footnote{A Cournot game is
where firms compete only on quantity, and sell all of the quantity at the
market clearing price.} The difference is that instead of being assigned a
market clearing price and selling everything they produce, in our model the
firms compete on price in a second subgame where they choose a price. This
subgame is necessary because the idiosyncratic tax can provide incentive for
firms to charge different prices. For example, a firm with low costs trying to
sell to non-local buyers would need to price undercut those customer's local
firm.

\subsection{Pricing Game}
Each firm $i$, having chosen a capacity $\hat q_i$, now chooses a price, $p_i$.
Consider two vending machines right beside one another selling the same soda,
with one machine selling soda at a cheaper price than the other. In this
subgame, buyers will use the cheaper machine first, but if the buyers demand
more soda than that machine's capacity, buyers will switch to the more
expensive machine when the cheaper one runs out of soda. Competition on price
in such a manner is called an Edgeworth game.

@@

Let the set of buyers be called $B$. Each buyer $b\in B$ buys a quantity $q_b =
\sum_{i\in S} q_{b,i}$, where $q_{b,i}$ is the amount buyer $b$ buys from the
firm $i$. The buyer's all have identical demand curves $p_\text{max} + q_b =
a$, where $a$ is the endowment they are given exogenously, and $p_\text{max}$
is the maximum price they pay (including tax). Each buyer tries to maximize
$q_b$ while keeping $p_\text{max} + q_b \le a$. In other words, they buy from
the cheapest available firm until they hit the demand curve and stop buying, or
the firm runs out of capacity and the buyer begins buying from the next
cheapest, repeating the process.

@@

\subsection{Profit Function}
Finally, firm $i$, having chosen quantity $\hat q_i$ and price $p_i$, and
having sold some or all of their capacity to buyers, will earn a profit:

@@

Also, each $q_{b,i}$ is chosen by the buyer $b$ so that they buy as much as
possible while staying under their demand curve $P+Q\leq a$; this is where
taxes affect the firm's profit.

@@

\subsection{Numerical Approximation}
Because the buyers have individualized preferences and the pricing game does
not necessarily have a `best' price choice, it is difficult to determine a
generalized, theoretical solution to the model laid out above. Instead, we
numerically determine the Nash equilibrium.\footnote{A Nash equilibrium
for a game is a set of choices, one for each player, such that upon seeing the
other firm's choices, could increase their payoff by changing strategies. In
our model, the players are firms, the payoff is profit, and the choices are
either capacity or price, depending on the subgame.}

To find the capacities and prices that the firms will choose, we begin by
discretizing the capacity and pricing choices so we only have a finite number
of possible combinations. Then, for every outcome of the capacity game, we find
the Nash equilibrium of the pricing game using the Python library Gambit
\cite{gambitPython}. Then, using the equilibrium prices, we calculate the
profits for every outcome of the capacity game. We finally use this information
to find the Nash equilibrium of the capacity game.

\section{Examples}
In Figure \ref{fig:fig3x2SingleTimestep} we can see the idiosyncratic tax
dampen Firm 1's attempts to become a monopoly power. In Rows 1 and 2, there is
no tax as $\gamma = 0$. While Firm 1 and 2 have an equal cost per unit, no firm
can become a monopoly. But if we reduce Firm 1's cost per unit by $1\%$, they
sell 14\% more than their competition, and earn 33\% more profit, which is out
of proportion to Firm 1's actual accomplishment, and can lead to
anti-competitive behavior by Firm 1.

In contrast, in Rows 3, 4, and 5, we let $\gamma = 1$. In Row 3 we can see a
mirror of Row 1, where both firms, faced with the same costs, produce the same
quantity, and make the same profit. But in row 4, where we give Firm 1 the same
advantage we game them in the row 2, they sell the same amount, and only make
12\% more profit than Firm 2, which implies that our tax protects Firm 2 from
small decrease in Firm 1's cost per unit. To increase their market share, Firm
1 needs to reduce their costs to 80\% of Firm 2, as shown in row 5.

The idiosyncratic tax thus dampens monopolistic tendencies of firms with slight
competitive advantages, while still ensuring that minor innovations are
profitable. But, if one firm is significantly better, as in Row 5, then that
firm will still dominate the market.

\begin{figure}[[p!htb] % p stands for 'entire page', '!htb' is to insert as close to code as possible
  \floatpagestyle{empty} % remove page number
  \centering
  \vspace{-2cm} % shift upwards to get more space for caption
  \includegraphics[width=\linewidth]{5x2plot.png}
  \caption{\small{Above are the model's results for a variety of
  conditions. The Rows 1 and 2 have $\gamma=0.0$, while Rows 3, 4, and 5 have
  $\gamma = 0.5$. In Rows 1 and 3, the firms have an equal cost per unit, sell
  equal quantity, and make equal profit. In Row 2, where $\gamma=0$, a 1\%
  difference in cost implies Firm 1 sells 14\% more quantity, and earns a 33\%
  larger profit.
  In Row 4, where $\gamma=.5$, Firm 1's 1\% difference in cost now results in
  an equal market share, but they still profit 11\% more. To get a larger
  market share, as in Row 5, Firm 1 must have a 20\% lower cost per unit. This
  implies that $\gamma$ dampens a firm's ability to gain market power from
  small innovations, without significantly discouraging innovation. 
  %[Check numbers after new gamma]
  }}
  \label{fig:fig3x2SingleTimestep}
\end{figure}

\section{Dynamics}
We run this model once for every timestep, then update the costs for the firms
between the timesteps. Every timestep we split a 5\% decrease in costs between
the firms according to the fraction of total profit the Firm makes. For
example, if Firm 1 earned a profit of 80, and Firm 2 earned a profit of 20,
then Firm 1's per unit cost would decrease by 4\%, and Firm 1's by 1\%.

\newpage
\printbibliography
\end{document}
